---
title: "mpg - r4ds data visualization exercise"
author: "Hafizhah Fadhilla"
date: "8/17/2021"
output: 
  html_document:
    toc: true
    toc_float: 
      collapse: false
      smooth_scroll: false
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gridExtra)
```


### Dataset

This dataset contains a subset of the fuel economy data that the EPA makes available on <https://fueleconomy.gov/>. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car. This dataset is inside the `ggplot2` of the `tidyverse` packages. 

```{r package, error=TRUE, message=FALSE}
data("mpg")

str(mpg)
```
The dataset contains 234 records and 11 variables. There are six categorical variable and five continuous variable.

1. `manufacturer`: manufacturer name 
2. `model`: model name
3. `displ`: engine displacement (litres)
4. `year`: year of manufacture
5. `cyl`: number of cylinders
6. `trans`: type of transmission
7. `drv`: the type of drive train (f = front-wheel drive, r = rear wheel drive, 4 = 4wd). front-wheel drive (FWD) means that the power from the engine is only delivered to the front wheels, on the contrary to rear-wheel drive (RWD). 4WD means the power from the engine is delivered to all 4 wheels all the time.
8. `cty` and `hwy`: measure car's fuel efficiency in miles per gallon (mpg) for city and highway cycles.
9. `fl`: fuel type
10. `class`: type of car

### 3.2 Exercise

> 1. Run the `ggplot(data = mpg)`. What do you see?

```{r ggplot_data}
ggplot(data = mpg)
```
It only showed blank canvas of the plot since no data mapped into x-axis and y-axis.


> 2. How many rows are in `mpg`? How many columns?

```{r}
dim(mpg)
```
The mpg dataset has 234 observations (rows) and 11 variables (columns).


> 3. What does the drv variable describe? Read the help for ?mpg to find out.

```{r drv variable}
?mpg 
```
It describes the type of drive train (f = front-wheel drive, r = rear wheel drive, 4 = 4wd). An alternative code to find the description is `help(mpg)`.


> 4. Make scatterplot of `hwy` vs `cyl`

```{r hwy_cyl plot}
ggplot(data = mpg) + 
  geom_point(mapping = aes(cyl, hwy)) +
  labs(title = "Highway MPG and Number of Cylinders")
```
The plot shows mpg tends to decrease as the number of cyl increases.


> 5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

```{r class_drv plot}
ggplot(data = mpg) + 
  geom_point(mapping = aes(drv, class)) +
  labs(title = "Drive Train and Types of Car")
```

```{r class_drv count}
mpg %>%
  count(class, drv)
```

`class` and `drv` are categorical variables. The count calculation resulted in 12 observations of combination `class` and `drv`. These unique combination of `class` and `drv` (not the resulting number of count calculation) are what plotted to the x and y axis, so plotting categorical variables into a scatterplot will not give valuable information. 


### 3.3 Exercise

> 1. What’s gone wrong with this code? Why are the points not blue?

```{r displ_hwy plot}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

The color of the points didn't changed because it's mapped into `aes`. To set an aesthetic manually, set the aesthetic by name as an argument of your geom function; e.g., it goes outside of aes().

> 2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg?

The categorical variables in `mpg` are: `manufacturer`, `model`, `trans`, `drv`, `fl`, and `class`; the continuous or numerical variables: `displ`, `year`, `cyl`, `cty`, and `hwy`.
To  see the types of variables in dataset, use `str()` or `glimpse()`.

```{r dataset structure}
str(mpg)

glimpse(mpg)
```
> 3. Map a continuous variable to `color`, `size`, and `shape`. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r mapping continuous variable and categorical variable, error=TRUE}

# mapping continuous variables in color, size, and shape
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy, color = displ))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = hwy, size = displ))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = hwy, shape = displ))

# mapping categorical variables in color, size, and shape
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy, color = fl))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = hwy, size = fl))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = hwy, shape = fl))

```

- When mapping categorical variable to `color`, it will create discrete color scale. When the numerical variable mapped into `color` aesthetic, it will display continuous color scale. 
- When a categorical variable mapped into `size`, R will display a warning. R needs ordered variable for `size` which is an ordered aesthetic mapping. 
- The `shape` aesthetic can be specified using an integer (between 0 and 25), a single character to specified the plotting symbol, an NA to draw nothing, or a mapping to discrete variable. R produces error when a continuous variable is mapped into `shape`.

> 4. What happens if you map the same variable to multiple aesthetics?

```{r multiple mapping varible into aes}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy, color = fl, size = fl, shape = fl))
```

It is not advisable to map the same variable into multiple aesthetic because the plot become ineffective to provide information to the reader.


> 5. What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

`stroke` is used to change thickness of the line of non-filled shapes.


> 6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = `displ` < 5)? Note, you’ll also need to specify x and y.

```{r aes mapping of non-variable}
p1 <- ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = hwy, color = displ < 5))

p2 <- ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = hwy, size = displ < 5))

grid.arrange(p1, p2, ncol = 2)
```

If we set condition to a variable that we use in aesthetic mapping, ggplot will produce a scale with logical value (`TRUE` or `FALSE`). If a value is set to a variable in aesthetic mapping, like `color = "blue"`, ggplot will create a color scale with that category.  


## 3.5 Exercise

> 1. What happens if you facet on a continuous variable?

```{r facet cont var, error=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~cty)
```

When a continuous variable is used to define a facet group, it will be converted into a categorical variable and ggplot will facet for each category.  


> 2. What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?

```{r facet_grid comparison}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```
The empty cells in `facet_grid(drv ~ cyl)` means that there is no data in the dataset that belongs to these facet groups (RWD with 4 and 5 cylinder; 4WD with 5 cylinder). The second plot also confirmed this by displying no data for 4WD with 5 cylinder and RWD with 4 and 5 cylinder.

> 3. What plots does the following code make? What does . do?

```{r}
p1 <- ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

p2 <- ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)

grid.arrange(p1, p2, ncol = 2)
```

A set of variables or expression is mapped in the `facets` argument of the `facet_wraps`. The formula can be written as `a ~ b` with `a` as row dimension and `b` as column dimension. `~ a` basically to facet the plot with row dimension, and `a ~ .` to facet the plot with column dimension.


> 4. Take the first faceted plot in this section:
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

```{r plot question}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

Since the faceted groups have their own panel, facetting may be useful to avoid overlapping data and observe the pattern of data distribution within and between groups. If there are too many the groups in the faceting variables, facetting become ineffective. When using facet, because the groups are not close together, small differences are harder to see, this is not the case for grouping with aesthetic.


> 5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

`nrow` and `ncol` arguments in `facet_wrap` are used to define the number of rows and columns for faceting. `nrow = 2` will create a plot with 2 rows facet.


> 6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

It will be easier to see the variable values on y-axis.


### 3.6 Exercise

> 1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

Line chart, boxplot, and histogram are types of geometry object in R. In ggplot2, to draw a line chart use `geom_line`, to create a boxplot chart use `geom_boxplot`. To produce a histogram plot use `geom_histogram`. 


> 2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r geom object question, echo = FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

This R code will produce a scatterplot with displacement engine (`displ`) on x-axis, highway mpg (`hwy`) on y-axis and aesthetic color from drive train (`drv`). The `geom_smooth()` function will create a trend line for the plot. In the code, `se = FALSE` so the confidence interval ribbon around the trend line will not be displayed. The aesthetic mappings are passed to `ggplot()`, so ggplot2 will treat these as global mappings that apply to each geom. As the result, ggplot2 will create trend lines for each group of `drv`.

```{r geom object answer prediction}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


> 3. What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

When `show.legend` set to `FALSE`,  the legend of the plot will be hided. If the `show.legend = FALSE` being removed, R will automatically display the legend created from the code. The author used it earlier in the chapter to show distinction of aesthetic mapping by using `group` and `color`.


> 4. What does the se argument to `geom_smooth()` do?

`se` argument is used to display the confidence interval ribbon around the smooth.

> 5. Will these two graphs look different? Why/why not?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

Both codes will produce identical plots. In the first code, data and aesthetic mapping put in the `ggplot()` function, so it acts as global mapping for the subsequent geoms. In the second code, data and aesthetic mapping were plotted in the geom objects. The first code is more convenient to write, because the same code doesn't need to be repeated.

> 6. Recreate the R code necessary to generate the following graphs.
> 
> Refer to [3.6.1 Exercise](https://r4ds.had.co.nz/data-visualisation.html#geometric-objects)

```{r recreate plots}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(aes(group = drv), se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv)) +
  geom_smooth(se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(aes(group = drv, linetype = drv), se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy, fill = drv)) +
  geom_point(shape = 21, color = "white", stroke = 2, size = 3)
```


### 3.7 Exercise

The diamonds dataset comes in ggplot2 and contains information about ~54,000 diamonds, including the `price`, `carat`, `color`, `clarity`, and `cut` of each diamond.

```{r diamonds}
#load the dataset
data("diamonds")
```

> 1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

Previous plot:

```{r stat summary plot}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

```{r pointrange plot}
ggplot(data = diamonds) + 
  geom_pointrange(aes(x = cut, y = depth),
                  stat = "summary",
                  fun.min = min,
                  fun.max = max,
                  fun = median
                  )

#Summary for diamonds cut data
diamonds_summary_cut <- diamonds %>% 
  group_by(cut) %>%
  summarise(median = median(depth),
            ymin = min(depth),
            ymax = max(depth)
            )

#Custom Plot 1
ggplot(data = diamonds_summary_cut, aes(x = cut, y = median)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax))

#Custom Plot 2
ggplot(data = diamonds_summary_cut, aes(x = cut, y = median)) +
  geom_point(size = 21) +
  geom_linerange(aes(ymin = ymin, ymax = ymax))
```

`stat_summary()` operates on unique x or y (it's `cut` category in diamonds data) and use default geom "pointrange". 


> 2. What does `geom_col()` do? How is it different to `geom_bar()`?

`geom_bar()` and `geom_col()` are function to create bar charts. `geom_bar` uses `stat_count()`, while `geom_col()` uses `stat_identity()`. In summary. `geom_bar(stat = "identity")` is same as `geom_col()`.


> 3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

```{r pair geom and stat, include=FALSE}
geom_object <- c(geom_bar(), geom_col(), geom_point(), geom_histogram(), geom_pointrange(), geom_area())

stat_function <- c(stat_count(), stat_identity(), stat_identity(), stat_bin(), stat_summary(), stat_density())
```

They act as an alternative to each other. Each pair will display the same result.

> 4. What variables does `stat_smooth()` compute? What parameters control its behaviour?

```{r stat_smooth}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  stat_smooth()

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth()
```
`stat_smooth()` and `geom_smooth()`are aliases. These function compute trend line to a scatter plot and apply smoothed method that accepts argument NULL or character vector, e.g. "lm", "glm" "gam", "loess" or a function.

> Q5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = after_stat(prop)))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = after_stat(prop)))
```


### 3.8 Exercise

> 1. What is the problem with this plot? How could you improve it?

```{r mpg plot 38}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

The points are overlapping. To improve this, use `geom_jitter()` or add a new argument `position = "jitter"`.

```{r mpg plot 38 fix}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()
```

> What parameters to `geom_jitter()` control the amount of jittering?

```{r geom_jitter control jitter}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter(width = 0.8, height = 0.8)
```
`width` and `height` arguments control the amount of vertical and horizontal jitter. The defaults is 40% of the resolution of the data.


> 3. Compare and contrast `geom_jitter()` with `geom_count()`.

```{r geom_jitter vs geom_count}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()
```

`geom_point()` maps individual point (observation) between two variables to the point area. If a dataset has 234 values, then each values will be plotted to the position of x and y. `geom_count()` is a variantion of `geom_point()` that counts the number of observations at each location, then maps the count to point area.


> What’s the default position adjustment for `geom_boxplot()`? Create a visualisation of the mpg dataset that demonstrates it.

```{r position geom_boxplot}
ggplot(mpg, aes(x = factor(cyl), y = cty)) +
  geom_boxplot()
  
```
The default position adjustment for`geom_boxplot()` is `"dodge2"`.


### 3.9 Exercise

> 1. Turn a stacked bar chart into a pie chart using `coord_polar()`.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity)) + 
  coord_flip() +
  coord_polar()
  
```


> 2. What does `labs()` do? Read the documentation.

To modify labels: title, subtitle, caption, tag, alt, alt_insight, and axis labels.


> 3. What’s the difference between coord_quickmap() and coord_map()?




# References
* [Understanding Your Drivetrain RWD, FWD, AWD](https://www.newroads.ca/blog/understanding-drivetrain-rwd-fwd-awd/).
* [Fundamentals of Fuel Consumption](https://www.nap.edu/read/12924/chapter/4).
* [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/).
* [R for Data Science](https://r4ds.had.co.nz/)